---
title: 'Database with R: PostgreSQL'
author: CXiao
date: '2018-05-30'
slug: database-with-r-postgresql
categories:
  - Data Science
  - Database
tags:
  - R
header:
  caption: ''
  image: ''
---

### Load library
```{r message = F, warning = F}
library(tidyverse)
library(RPostgres)
library(DBI)
```

### Connect to **PostgreSQL**  
### Connect Method 1 with `RPostgres`
```r
con1 <- DBI::dbConnect(RPostgres::Postgres(), 
                      dbname = "testdb1",
                      password = rstudioapi::askForPassword("Database password"))

dbListTables(con1)
#dbWriteTable(con1, "mtcars", mtcars)
#dbWriteTable(con1, "flights", nycflights13::flights)
```
`rstudioapi::askForPassword` requires the password input

* In the command line (or block-run), there will be popout dialog to input password,  
* In the knit mode, `render` the RMarkdown file as follows:  
      `rmarkdown::render("MyDocument.Rmd", params = "ask")`  
      See [Parameter User Interfaces](https://rmarkdown.rstudio.com/developer_parameterized_reports.html#parameter_user_interfaces)
* In the `blogdown`, `server_site` will halt at the above block.  

### Connect Method 2 with `RPostgreSQL`
```r
library(RPostgreSQL)
drv <- dbDriver("PostgreSQL")
con2 <- dbConnect(drv, 
                 host = "localhost",
                 dbname = "testdb1",
                 user = "cxiao",
                 password = rstudioapi::askForPassword("Database password"))
dbListTables(con2)
```

### Connect Method 3 with `RPostgres` and `config.yml`
```{r}
Sys.setenv(R_CONFIG_ACTIVE = "AWS")
dw <- config::get(file = "config.yml", "dw")

conAWS1 <- dbConnect(RPostgres::Postgres(),
                     dbname   = dw$database,
                     user      = dw$uid,
                     password = dw$pwd,
                     host     = dw$server,
                     port     = 5432)
dbListTables(conAWS1)
```

### Connect Method 4 with `odbc` and `config.yml` 
```{r}
conAWS2 <- dbConnect(odbc::odbc(),
                     driver   = dw$driver,
                     database = dw$database,
                     uid      = dw$uid,
                     pwd      = dw$pwd,
                     server   = dw$server,
                     port     = 5432)

dbListTables(conAWS1)
```

#### SQL ETL
```{sql connection=conAWS1}
SELECT * FROM mtcars LIMIT 10
```

#### SQL in R
```{r}
dbGetQuery(conAWS1, "SELECT * FROM mtcars LIMIT 10")
```

#### ddplyr 1
```{r}
mtcars_db <- tbl(conAWS1, "mtcars")

top2 <- mtcars_db %>% 
  group_by(cyl) %>%
# slice not working for database 
  filter(between(row_number(), 1, 2)) 
```

#### ddplyr 2
```{r}
flights_db <- tbl(conAWS1, "flights")

flights_db

flights_db %>% 
  tally()
```


