---
title: 'Time Series Analysis I: date and time '
author: ''
date: '2018-04-20'
slug: time-series-analysis-i-date-and-time
categories:
  - Data Science
  - Time Series
tags:
  - R
header:
  caption: ''
  image: ''
---

## `Date` class from `base::Date` and `lubridate`  
```{r}
d1 <- Sys.Date()
d1; class(d1)

dates <- c("02/27/92", "02/27/92", "01/14/92")
dd    <- as.Date(dates, "%m/%d/%y")
dd; class(dd)

d2 <- lubridate::as_date("2018-07-05")
d2; class(d2)
```

## `POSIX*` class from `base::as.POSIX*`  

`POSIXt` types, `POSIXct` and `POSIXlt`:  

* `ct`, calendar time, stores the number of seconds since the origin   
* `lt`, local time, keeps the date as a list of time attributes (such as
”hour” and ”mon”).   
   `unclass(as.POSIXct(Sys.time()))` vs `unclass(as.POSIXlt(Sys.time()))`

```{r}
dt1 <- as.POSIXct(Sys.time())
dt1; class(dt1)

dt2 <- lubridate::as_datetime("2018-07-05 15:30:21")
dt2; class(dt2)
```


## `timeDate` class from `timeDate`
```{r}
dts <- c("1989-09-28", "2001-01-15", "2004-08-30", "1990-02-09")
tms <- c(  "23:12:55",   "10:34:02",   "08:30:00",   "11:18:23")
tD1  <- timeDate::timeDate(dts, format = "%Y-%m-%d", FinCenter = "GMT" )
class(tD1)
```

## `dates`, `times` and `chron` class from `chron`
```{r}
dts <- chron::dates(c("02/27/92", "02/27/92", "01/14/92"))
dts; class(dts)

dts <- chron::dates(c("02/27/92", "02/27/92", "01/14/92"))
tms <- chron::times(c("23:03:20", "22:29:56", "01:03:30"))
tms; class(tms)
# [1] 23:03:20 22:29:56 01:03:30 18:21:03 16:56:26
x <- chron::chron(dates = dts, times = tms)
x; class(x)
```