---
title: 'Comparison: sort, order and arrange'
author: Chuliang Xiao
date: '2018-03-28'
slug: comparison-sort-order-and-arrange
categories:
  - Comparison
  - Data Science
tags:
  - R
header:
  caption: ''
  image: ''
---
* **vector** (or factor)  
---  
```{r}
(x <- swiss$Education[1:20])
# sort the vector
sort(x)
# partial sort ing
sort(x, partial = c(10, 15))
```
Partial sorting in R is different with that in [Wikipedia](http://en.wikipedia.org/wiki/Partial_sorting). See [Stack Overflow](https://stackoverflow.com/a/23579512)

* **data.frame**  
--- 
1. order
```{r}
head(mtcars)
# sort by mpg
Rank1 <- mtcars[order(mtcars$mpg), ]
head(Rank1)
# sort by cyl, then by hp
Rank2 <- mtcars[order(mtcars$cyl, mtcars$hp), ]
head(Rank2)
# reverse sort
Rank3 <- mtcars[order(-mtcars$cyl, mtcars$hp), ]
head(Rank3)
```
If data.frame mtcars is `attach()`, then
```{r}
attach(mtcars)
mpgRank <- mtcars[order(mpg), ]
head(mpgRank)
```
2. dplyr::arrange
```{r message = F, error = F, warning = F}
library(dplyr)
```

```{r warning = F}
Rank4 <- arrange(mtcars, mpg)
head(Rank4)

Rank5 <- arrange(mtcars, cyl, hp)
head(Rank5)

Rank6 <- arrange(mtcars, -cyl, hp)
head(Rank6)
```

3. difference between order and arrange  
    + row names: `order` keep the row names of mtcars, while not in `dplyr::arrange`  
    + efficency: `dplyr::arrange` tends to be faster when there are lots of rows.  


Refer to  
1. [Cookbook for R](http://www.cookbook-r.com/Manipulating_data/Sorting/)  
2. [Quick-R](https://www.statmethods.net/management/sorting.html)  