---
title: 'Comparison: transform, within and mutate'
author: Chuliang Xiao
date: '2018-03-25'
slug: comparison-transform-within-and-mutate
categories:
  - Comparison
  - Data Science
tags:
  - R
header:
  caption: ''
  image: ''
---
Here is another comparison between two basic functions `transform` and `within`, and a tidyverse function `dplyr::mutate`. They all can be used for data munipulation, adding a new column to a data.frame.  
```{r}
head(mtcars)  
```

### One variable
Calculate horse power per cylinder (`hp/cyl`) and add it to `mtcars`
```{r warning = F, message = F}
tmp1  <- transform(mtcars, hpc = round(hp / cyl))
head(tmp1)
tmp2 <- within(mtcars, {hpc <- round(hp / cyl)})
# tmp1 and tmp2 are identicial
head(tmp2)
library(dplyr)
tmp3 <- mutate(mtcars, hpc = round(hp / cyl))
# tmp3 is ALMOST the same as tmp1 and tmp2, except that row names are missed
head(tmp3)
```
### two variables
A variable created by `transform` process cannot be used[^1].  
The example below fails.

`tmp1  <- transform(mtcars, hpc = round(hp / cyl), hpc1 = hpc + 1)`

While `within` and `dplyr::mutate` are OK[^2].  
```{r warning = F, message = F}
tmp2 <- within(mtcars, {hpc  <- round(hp / cyl)
                        hpc1 <- hpc + 1 })
head(tmp2)

tmp3 <- mutate(mtcars, hpc = round(hp / cyl), hpc1 = hpc + 1)
head(tmp3)
```


[^1]: What is the difference between `transform` and `within`
[Stack Overflow](https://stackoverflow.com/a/22417775)
[^2]: [Data manipulation using transform(), within(), or plyr::mutate()](http://rstudio-pubs-static.s3.amazonaws.com/1053_c089bd9190624b37b1180b91d2776e10.html)