---
title: 'Time Series Analysis II: zoo'
author: ''
date: '2018-05-10'
slug: time-series-analysis-ii-zoo
categories:
  - Data Science
  - Time Series
tags:
  - R
header:
  caption: ''
  image: ''
---

[**zoo**](https://cran.r-project.org/web/packages/zoo/): S3 Infrastructure for Regular and Irregular Time Series (Z's Ordered Observations)  

> An S3 class with methods for totally ordered indexed observations. It is particularly aimed at irregular time series of numeric vectors/matrices and factors. zoo's key design goals are independence of a particular index/date/time class and consistency with ts and base R by providing methods to extend standard generics  


### Base Time-Series Objects `stats::ts` 
Credit: _Time Series Analysis in R Part 1: The Time Series Object_ by [DataSciencePlus](https://datascienceplus.com/time-series-analysis-in-r-part-1-the-time-series-object/)  

> From the base `ts` objects to a whole host of other packages like `xts`, `zoo`, `TTR`, `forecast`, `quantmod` and `tidyquant`, R has a large infrastructure supporting time series analysis.  

#### Monthly/Quaterly TS
```{r}
ts1 <- ts(1:10, frequency = 12, start = c(1959, 2)) # Feburary of 1959
ts1; class(ts1)

ts2 <- ts(1:10, frequency = 4, start = c(1959, 2)) # 2nd Quarter of 1959
ts2
```

#### Period TS
```{r}
ts3 <- ts(1:10, frequency = 7, start = c(2012, 2))
print(ts3, calendar = TRUE)
```

#### Plot TS
```r
gnp <- ts(cumsum(1 + round(rnorm(100), 2)),
          start = c(1954, 7), frequency = 12)
plot(gnp) # using 'plot.ts' for time-series plot
```
### `zoo` class 
#### Read vector with a time index
```{r warning=F, message=F}
library(zoo)
f1 <- system.file('doc/demo1.txt', package = 'zoo')
system(paste0("head ", f1))
inrusd <- read.zoo(f1, sep = "|", format="%d %b %Y")
class(inrusd)
str(inrusd)
head(inrusd)
```

#### Read matrix with a time index
```{r}
f2 <- system.file('doc/demo2.txt', package = 'zoo')
tmp <- read.table(f2, sep = ",")
z <- zoo(tmp[, 3:4], as.Date(as.character(tmp[, 2]), format="%d %b %Y"))
colnames(z) <- c("Nifty", "Junior")

head(z)
```

#### Convert back to matrix
```{r}
plain1 <- coredata(z)
head(plain1)

# with rownames
plain2 <- as.matrix(z)
head(plain2)
```

#### Selecting (subsetting)
```{r}
window(z, start = as.Date("2005-02-15"), end = as.Date("2005-02-28"))
z[as.Date("2005-03-10")]
```


#### Missing Value
```r
# only rows with data from both x and y are included in the output
m1 <- merge(inrusd, z, all = FALSE)
plot(m1)
```
```{r}
# then extra rows will be added to the output with missing values
m2 <- merge(inrusd, z)
plot(m2)

m_approx <- na.approx(m2)# Replaced by linear interpolation via approx 
m_locf   <- na.locf(m2)  # Last Observation Carried Forward
```


#### Plot
```{r fig.width=10, fig.height=5}
library(ggplot2)
x.Date  <- as.Date(paste(2003, 02, c(1, 3, 7, 9, 14), sep = "-"))
x       <- zoo(rnorm(5), x.Date)
xlow    <- x - runif(5)
xhigh   <- x + runif(5)
#z       <- cbind(x, xlow, xhigh)

## univariate plotting
## calling ggplot2.zoo
## autoplot(x)

## broom zoo to data.frame
ggplot(aes(x = index, y = value), data = broom::tidy(x)) +
  geom_line() + xlab("Index") + ylab("x")
```