---
title: 'R Spatial Data Analysis 1: from Data to Spatial Data'
author: CXiao
date: '2018-03-31'
slug: r-spatial-data-analysis-1-from-data-to-spatial-data
categories:
  - Data Science
  - Spatial
tags: 
  - R
header:
  caption: ''
  image: ''
---
As much as R is popular in data analysis, R becomes more and more favored in geospatial analysis and visualization.  

### Data and Plots  

#### basic plot  
```{r message = F}
attach(mtcars)
par(mfrow = c(1,2))
plot(mpg, wt, main = "wt vs. mpg")
hist(mpg)
```

### ggplot2 
```{r message = F}
library(ggplot2)
p1  <- ggplot(mtcars, aes(x = mpg, y = wt)) + 
  geom_point() +
  ggtitle("wt vs. mpg") +
  theme(plot.title = element_text(hjust = 0.5))
p2  <- ggplot(mtcars, aes(mpg)) +
  geom_histogram(binwidth = 5, color = 'black', fill = 'white') +
  ggtitle("mpg Histogram") +
  theme(plot.title = element_text(hjust = 0.5))  
library(grid)
library(gridExtra)
grid.arrange(p1, p2, ncol = 2)
```


#### Spatial Data and Plots

### Shape objects

> **shape objects** as objects from the class `Spatial` or `Raster`, respectively from the `sp` and the `raster` package. The supported subclasses are[^1]

|   	| Without data	| With data   	|
|---	|---	|---	|
|Polygons | SpatialPolygons	| SpatialPolygonsDataFrame|
|Points 	| SpatialPoints  	| SpatialPointsDataFrame	|
|Lines   	| SpatialLines  	| SpatialLinesDataFrame  	|
|Raster  	| SpatialGrid  		| SpatialGridDataFrame  	|
|Raster		| SpatialPixels  	| SpatialPixelsDataFrame 	|
|Raster		| 					| RasterLayer				|
|Raster		| 					| RasterBrick				|
|Raster		| 					| RasterStack				|

### Reading
* Tutorial: [Creating maps in R](https://github.com/Robinlovelace/Creating-maps-in-R) by Robin Lovelance and James Cheshire
* website [r-spaital.org](https://www.r-spatial.org) by Edzer Pebesma  
* CRAN Task View: [Analysis of Spatial Data](https://CRAN.R-project.org/view=Spatial) by Roger Bivand 

### Packages
* `rgdal`: R's interface to Geospatial Data Abstraction Library (GDAL)  
    + `rgdal::readOGR`, to read OGR vector maps (e.g. shapfile) into Spatial objects  
* `rgeos`: R's interface to the powerful vector processing library geos  
* `maptools`: provides various mapping functions  
* `tmap`: a new packages for rapidly creating beautiful maps  
    + An extenstion wrap of the basic `plot`
    + `qtm`, quick thematic map plot  
    + `tm_shape`, `tm_fill`
* `ggmap`: extends the plotting package `ggplot2` for maps 
    + add a variaty of basemaps
* `leaflet`: interactive map from its [JavaScript](http://leafletjs.com/)  
* `mapview`: interactive viewing spatial data based on `leaflet`
    + Personally I like `mapview`'s popup function, see [tutorial](https://r-spatial.github.io/mapview/)   
    + It spports `sp`, `raster`, `sf` and [`satellite`](https://cran.r-project.org/package=satellite)

##### maptools::readShapeSpatial  

> The use of `maptools::readShapeSpatial` has been **deprecated**. Use `rgdal::readOGR` and `sf::st_read` instead.

### Beyond `sp` package
* `sf` [simple feature package](../r-spatial-data-analysis-3-simple-features)  
* `star` [spatiotemporal didy arrays for R](../r-spatial-data-analysis-4-stars)  

[^1]: [tmap in a nuthell](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-nutshell.html)